---
output:
  rmarkdown::github_document:
    html_preview: no
---

<!--- README.md is generated from README.Rmd. Please edit that file -->

# Survey Value of Information

[![lifecycle](https://img.shields.io/badge/Lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

```{r, include = FALSE}
knitr::opts_chunk$set(fig.path = "man/figures/README-", fig.align = "center")
```

```{r, include = FALSE}
devtools::load_all()
h = 3.5
w = 3.5
ow = "400"
```

TODO OVERVIEW TODO.

## Installation

The *surveyvoi R* package has several dependencies that must first be installed before you can install this package. Specifically, you will first need to install the Intel Math Kernel Library [(Intel MKL)](https://software.intel.com/en-us/mkl). For instructions on installing the Intel Math Kernel Library, please refer to the following bash commands (based on a [blog post by Dirk Eddelbuettel](http://dirk.eddelbuettel.com/blog/2018/04/15/)):

```{bash, eval = FALSE}
cd /tmp
wget https://apt.repos.intel.com/intel-gpg-keys/GPG-PUB-KEY-INTEL-SW-PRODUCTS-2019.PUB
sudo apt-key add GPG-PUB-KEY-INTEL-SW-PRODUCTS-2019.PUB
sudo sh -c 'echo deb https://apt.repos.intel.com/mkl all main > /etc/apt/sources.list.d/intel-mkl.list'
sudo apt-get -o Acquire::http::proxy=false update
sudo apt-get -o Acquire::http::proxy=false install intel-mkl-64bit-2020.0-088
```

You will then need to add the following commands to your `~./bashrc` file:

```
export MKLROOT="/opt/intel/mkl"
export LD_LIBRARY_PATH="${LD_LIBRARY_PATH}:${MKLROOT}/lib/intel64_lin"
```

Finally, after installing the dependencies, you can install the package with the following _R_ commands.

```{r, eval = FALSE}
if (!require(remotes))
  install.packages("remotes")
remotes::install_github("jeffreyhanson/surveyvoi")
```


## Usage

Just don't - it's still early in development and isn't ready yet.

## Citation

```{r, echo = FALSE, result = "asis", comment = ""}
citation("surveyvoi")
```
