---
output:
  rmarkdown::github_document:
    html_preview: no
---

<!--- README.md is generated from README.Rmd. Please edit that file -->

# Survey Value of Information

[![lifecycle](https://img.shields.io/badge/Lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

```{r, include = FALSE}
knitr::opts_chunk$set(fig.path = "man/figures/README-", fig.align = "center")
```

```{r, include = FALSE}
devtools::load_all()
h = 3.5
w = 3.5
ow = "400"
```

The _surveyvoi_ package is a decision support tool for prioritizing sites for ecological surveys based on their potential to improve conservation plans for managing biodiversity (e.g. plans for establishing protected areas). Given a set of sites that could potentially be acquired for conservation management -- wherein some sites have previously been surveyed and other sites have not -- this package provides functionality to identify which sites should be surveyed because doing so would likely lead to superior conservation management plans. It uses value of information analyses to measure return on investment for competing survey schemes. Furthermore, by directly maximizing the expected value of sample information, optimal survey schemes can also be identified.

## Installation

The _surveyvoi R_ package is still in an experimental phase of the development cycle and so it is not yet available on [Comprehensive R Archive Network (CRAN)](https://cran.r-project.org/). To install latest developmental version from [GitHub](https://github.com/jeffreyhanson/surveyvoi), please use the following R code. Note that developmental versions may contain coding errors.

```{r, eval = FALSE}
if (!require(remotes)) install.packages("remotes")
remotes::install_github("jeffreyhanson/surveyvoi")
```

## Usage

TODO.

## Citation

This package is still early in development and should not be used yet.
