---
output:
  rmarkdown::github_document:
    html_preview: no
---

<!--- README.md is generated from README.Rmd. Please edit that file -->

# Survey Value of Information

[![lifecycle](https://img.shields.io/badge/Lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

```{r, include = FALSE}
knitr::opts_chunk$set(fig.path = "man/figures/README-", fig.align = "center")
```

```{r, include = FALSE}
devtools::load_all()
h = 3.5
w = 3.5
ow = "400"
```

TODO OVERVIEW TODO.

## Installation

The latest version of the _surveyvoi R_ package can be installed using the following _R_ code. Note that before you can install this _R_ package, you will first need to install the [Gurobi optimization suite](http://www.gurobi.com/) and [_gurobi R_ package](https://www.gurobi.com/documentation/8.1/refman/r_api_overview.html). For instructions on installing _Gurobi_ software, please refer to [this installation guide](https://CRAN.R-project.org/package=prioritizr/vignettes/gurobi_installation.html).

```{r, eval = FALSE}
if (!require(remotes))
  install.packages("remotes")
remotes::install_github("jeffreyhanson/surveyvoi")
```

## Usage

Just don't - it's still early in development and isn't ready yet.

## Citation

```{r, echo = FALSE, result = "asis", comment = ""}
citation("surveyvoi")
```
